---
title: "Regression Models Course Project"
author: "Patrick"
date: "1/25/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# Coursera Data Science Course 7 Regression Models

## Exploratory analysis
```{r, include=FALSE}

library(ggplot2)
data(mtcars)
head(mtcars) ## note am = transmission (0 - automatic, 1 - manual)
summary(mtcars)

```

## Regression Analysis
```{r}
aggregate(mpg~am, data = mtcars, mean)
```
So far we can see that automatic cars have a lower mpg than manual cars.

Now lets find the variables that are highly correlated with mpg. We will use a 
step function to run miltiple linear models to build our multiple regression model.
```{r Appendix1, include=FALSE}

## creates linear models with mpg and remaining variables
lm1 <- lm(mpg ~ ., data = mtcars) 
m2 <- step(lm1, direction = "both") 
summary(m2) 
```


From what we can see above wt, qsec, and am produce the best fit model.
the adjusted r-squared also tells us that 83% of the variability is explained
by the model.

We can now compare the base model with the model produced by our step function
```{r}

m0 <- lm(mpg ~ am, data = mtcars)
anova(m0, m2)

```

The results above indicate that there is a significant contribution from wt
and qsec to the accuracy of the model.

# Residuals
Next we can check the residuals (Appendix fig. 2)
#### Summary
* residuals vs fitted - Points for most part randomly scattered so conditions seem to be indipendant.
* normal Q-Q -e points for the most part lie on the line indicatiing a normal distribution.
* Scale-Location - points seem to be scattered normaly around line indication constant variance.
* Residuals vs Leverage - From the graph we can see a few outliers

# Hypothesis testing
Lastly can perform a t.test do determine if the differences between a automatic
or manual transmission are significant.

```{r}
t.test(mpg ~ am, data = mtcars)
```

from the t test results we can see that the difference between cars with a automatic transmission
compared to a manual transmission is significant.

## Executive Summary

Based on the analysis conducted in this project we can conclude that cars with a Manual Transmission get 2.9358 more mpg per gallon adjusted for weight and quarter-mile time; a subsequent t-test showed the difference to be statistically significant.

## Appendix

### fig 1. Pairs PLOT denoting relationships for mtcars data table.

```{r}
## scatterplot matrix to visualize relationships
pairs(mpg~., data = mtcars, panel = panel.smooth, col = "9")

```

### fig 2. Residuals plot
```{r}

par(mfrow = c(2,2))
plot(m2)

```

### fig 3. Boxplot of mpg by transmission
```{r}

boxplot(mpg ~ am, data = mtcars, ylab = "mpg", xlab = "Transmission")

```














